---
title: "NumericData"
author: "Mia Fitzgerald"
date: "3/24/2026"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(Lock5Data)
library(tidyverse)
library(knitr)
data(StudentSurvey)

```


### Top 6 of New BMI Data


```{r}
student_health <- StudentSurvey %>%
  mutate(BMI = 701 * Weight/Height^2) %>%
  mutate(BMI = round(BMI, 1))

student_health %>%
  select(Sex, Weight, Height, BMI) %>%
  head(6) %>%
  kable() 

```


I mutated the data to make and round the data for a new BMI column. Then I stored that in a new dataset incase I want to use that column again. Then I select the only 4 columns I want to display for the first 6 instances in a table. 
^[Question 1 complete]


### Binning for Weight Status 


```{r}
student_health %>%
  mutate(Status = cut(BMI, c(-Inf, 18.5, 24.9, 29.9, Inf), labels = c("Underweight", "Healthy Weight", "Overweight", "Obese"))) %>%
  head(6) %>%
  kable()

```


I take the new dataset I made to to bin the BMI values and make a new column that renames the values into categorical health meanings. I display the first 6 instances in a table. 
^[Question 2 complete]


### Highest and Lowest BMIs for Male and Female Students


```{r}
student_health %>%
  group_by(Sex) %>%
  summarize(min_BMI = min(BMI, na.rm=TRUE), max_BMI = max(BMI, na.rm=TRUE)) %>%
  kable()

```


I take the same BMI dataset to group by sex in order to summarize operations the reflects both sex's. I got the minimum and maximum BMI's of each sex and put it in a table. 
^[Question 3 complete]


### Greater of Each Student’s Verbal and Math SAT Scores


```{r}
StudentSurvey %>%
  mutate(Max_SAT= pmax(VerbalSAT, MathSAT, na.rm=TRUE)) %>%
  select(VerbalSAT, MathSAT, Max_SAT) %>%
  head(6) %>%
  kable()
  
```


Taking the original dataset, I use pmax to compare the two SAT scores and make a new column with the greater score. Then I select the columns that helped make it, including itself. And displayed the first 6 and put it in a table. 
^[Question 4 complete]


### Z-Scores for Students’ Total SAT Scores


```{r}
StudentSurvey %>%
  mutate(TotalSAT = MathSAT + VerbalSAT, 
         Z_score = (TotalSAT - mean(TotalSAT)) / sd(TotalSAT)) %>%
  select(VerbalSAT, MathSAT, TotalSAT, Z_score) %>%
  head(6) %>%
  kable()

```


^[Question 5 complete]


